Computational Prediction of Boundaries of 3D Genomic Domains in Class Imbalance

### 2019-06-06 Manuscript/manuscript.Rmd
intro edits

The advent of various genome-wide sequencing technologies, such as high-throughput conformation capture (notably Hi-C), have revealed how the spatial organization of the human genome may affect genomic regulation [@lieberman2009comprehensive]. Analyses have shown that the genome is folded into distinct compartments termed topologically associating domains (TADs) [@dixon2012topological]. Approximately 60-80% of TADs remain stable across cell types [@Schmitt2016aa], suggesting that the boundaries of TADs may play a role in restricting the function of certain genomic elements such as enhancers, thereby impacting the transcription of genes. Furthermore, studies have shown that changing the 3D structure of the DNA, causing disruptions in these domains, can lead to adverse outcomes and diseases like cancer [@hnisz2016activation; @Flavahan2016; @song2017ctcf]. Therefore, it has become increasingly important to be able to identify the key molecular drivers of the formation of TAD boundaries in order to further our understanding of the human genome.

The distinct patterns of epigenomic marks, binding peaks of different factors, and actively transcribed genes prompt the use of computational approaches to identify which epigenomic features are most predictive of TAD boundaries [??? REFS? Carefully select from my paper notes]. Evidence suggests that regulatory elements and genes tend to interact more frequently within the same TAD [@symmons2014functional]. [??? This sentence and reference belong to the previous paragraph, think and move] It has been established that the zinc finger protein, CTCF, is associated with TAD boundaries and is an architectural protein responsible for the organization of the genome within the nucleus [@ghirlando2016ctcf; @ong2014ctcf]. CTCF along with other transcription factors, including SMC1, SMC3, and RAD21 <!--have been shown to -->colocalize <!--together [??? Note removal of "water" words]-->near TAD boundaries, leading to the recruitment of a cohesin ring [@sanborn2015chromatin; @fudenberg2016formation; @alipour2012self]. Together, these components are hypothesized to be what causes loops in the DNA, contributing to higher order chromatin organization. Other binding proteins have been shown to mediate enhancer-promoter interactions, including the transcription factors YY1 and ZNF143, thereby influencing the delineation of topological domains [@weintraub2017yy1; @gonzalez2017znf143; @heidari2014genome]. Even more so, loci within a domain have been shown to interact more frequently with one another, exhibiting similar patterns of chromatin modifications, and long-range contact patterns [@rao20143d]. [??? This sentence and reference belong to the previous paragraph, think and move] However, the growing number of annotated genomic features and the methods used to identify the ones associated with TAD boundaries still leaves much to be desired. [??? This is very weak and undefined. You started with mentioning prediction - finish with unsolved problems with it.]

>>>>

The advent of various genome-wide sequencing technologies, such as high-throughput conformation capture (notably Hi-C), have revealed how the spatial organization of the human genome may affect genomic regulation [@lieberman2009comprehensive]. Analyses have shown that the genome is folded into distinct compartments termed topologically associating domains (TADs) [@dixon2012topological]. Given the high degree of conservation between cell types and species, it has been proposed that TADs represent a fundamental unit of organization, giving the genome its architecture [@rocha2015breaking; @dixon2012topological]. Likewise, Evidence suggests that the boundaries that delineate TADs may play a role in restricting the function of certain genomic elements such as enhancers, thereby impacting the transcription of genes [@symmons2014functional]. Even more so, loci within a domain have been shown to interact more frequently with one another, exhibiting similar patterns of chromatin modifications, and long-range contact patterns [@rao20143d; @symmons2014functional] [??? This sentence and reference belong to the previous paragraph, think and move; !!!combined and moved]. Furthermore, studies have shown that changing the 3D structure of the DNA, causing disruptions in these domains, can lead to adverse outcomes and diseases like cancer [@hnisz2016activation; @Flavahan2016; @song2017ctcf]. Therefore, it has become increasingly important to be able to identify the key molecular drivers of the formation of TAD boundaries in order to further our understanding of the human genome.

The distinct patterns of epigenomic marks, binding peaks of different factors, and actively transcribed genes prompt the use of computational approaches to identify which epigenomic features are most predictive of TAD boundaries [dixon2015chromatin; @di2017novo] [??? REFS? Carefully select from my paper notes; !!!added refs]. It has been established that the zinc finger protein, CTCF, is associated with TAD boundaries and is an architectural protein responsible for the organization of the genome within the nucleus [@ghirlando2016ctcf; @ong2014ctcf]. CTCF along with other transcription factors, including SMC1, SMC3, and RAD21 <!--have been shown to -->colocalize <!--together [??? Note removal of "water" words]-->near TAD boundaries, leading to the recruitment of a cohesin ring [@sanborn2015chromatin; @fudenberg2016formation; @alipour2012self]. Together, these components are hypothesized to be what causes loops in the DNA, contributing to higher order chromatin organization. Other binding proteins have been shown to mediate enhancer-promoter interactions, including the transcription factors YY1 and ZNF143, thereby influencing the delineation of topological domains [@weintraub2017yy1; @gonzalez2017znf143; @heidari2014genome]. However, the growing number of annotated genomic features and the methods used to identify the ones associated with TAD boundaries still leaves much to be desired. [??? This is very weak and undefined. You started with mentioning prediction - finish with unsolved problems with it.]
